---
title: "Midterm"
author: "Zihan Yu"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(readxl)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data
The data set contains weight records of three pet dogs – Simone, Gagne, and Raisin. It has 6 variables: DATE, Raisin, Simone, Gagne, std and note, the DATE contains the date when the weight is recorded, the second to the fourth columns contains the weight of the dogs in Lbs and Oz, each column has two values. The std columns contains the weight of a standard object, it also has two values in Lbs and Oz.

First, reading the data from excel, take a glimpse of the dataset.

```{r}
dat <- read_excel("dogweights_07sept2020.xlsx", skip = 1)
colnames(dat)[6] <- "note"
head(dat)
```
The data is not in tidy format, since the DATE is ircorrectly formated, and the 2-4 columns have multiple values, so we should do some data cleaning before exploring. The notes only occured in a few days, too many NA values occured in this column, it takes up too many spaces, we can extract the note into a new table.

First, take use of the as.Date() function, convert the DATE from character to Date type.
```{r}
dat$DATE <- as.Date(as.numeric(dat$DATE), origin = "1899-12-30")
head(dat)
```

The weight contains two values in different unit, here I will only keep the weight in Lbs.

```{r}
getWeight <- function(x){
  if(is.na(x)){
    return(NA)
  }
  tokens <- unlist(strsplit(x, " "))
  as.numeric(tokens[1])
}
dat$Raisin <- sapply(dat$Raisin, getWeight)
dat$Simone <- sapply(dat$Simone, getWeight)
dat$Gagne <- sapply(dat$Gagne, getWeight)
dat$std <- sapply(dat$std, getWeight)
```

Get the summary statisitc of the cleaned dataset.
```{r}
summary(dat)
```
There are some outliers in the Raisin column, some values are extreamly larger, higher than 1800, and the minimum value is 7, which is impoissible. If we look up the original table, we can found the correct value is should be 17 and 18.

Fix the data by repalce the wrong value with the correct weight.
```{r}
dat$Raisin[dat$Raisin==7] <- 15
dat$Raisin[dat$Raisin>1800] <- 18
```

Check again, now all the weights are in a valid range.
```{r}
summary(dat)
```
Besides, the dataset is in wide format, yuu can recognize data in wide format by the fact that columns generally represent groups, as in our data, Raisin, Simone, Gagne is a group, all the three columns are the weights of a group, so it is in wide foratm we should convert it to long format, we can take use of the gather() in function in the tidyverse packages.
```{r}
dat <- dat %>% gather(key=Pet, value=Weight, Raisin:Gagne)
```

Create two dataset, one contains the date and weight, the other contains the date the note. Save the data into csv. In the first dataset, remove any row contains the NA value.
```{r}
dat1 <- dat[, c(1,4,5)]
dat1 <- na.omit(dat1)
write.csv(dat1, "dat1.csv", row.names = F)
dat2 <- dat[, c(1, 3)]
dat2 <- na.omit(dat2)
write.csv(dat2, "dat2.csv", row.names = F)
```





